\documentclass[b5paper,12pt,oneside,openright]{memoir}
\usepackage{amsmath, amsthm, amssymb}
\usepackage{enumitem}
\usepackage{color}

\setlength{\uppermargin}{22mm}
\setlength{\spinemargin}{22mm}
\setlength{\textheight}{206mm}
\setlength{\textwidth}{132mm}
\setlength{\oddsidemargin}{0mm}
\fixthelayout

\DeclareMathOperator{\sgn}{sgn}
\newcommand{\jacobi}[2]{\left(\frac{#1}{#2}\right)}

\begin{document}
\section{The relation between $p\mid x^2+ny^2$ and quadratic reciprocity}
Suppose $x^2 + ny^2 = kp$ for some $k$. Assume there is a common divisor of $y$ and $p$ which is prime; since $p$ is prime, it must be $p$.
Then $p\mid x^2$ which implies $p\mid x$, i.e. $p$ is a common divisor of $x$ and $y$.
This tells us $\gcd(x,y)=1 \implies \gcd(y,p)=1$.

In particular, $\gcd(x,y)=1\implies y$ is invertible modulo $p$ and
\[(xy^{-1})^2\equiv -n\bmod p,\]
so $-n$ is a quadratic residue modulo $p$.

On the other hand, if $-n\equiv x^2\bmod p$ for some $x$, $p \mid x^2+n\cdot 1^2$ (and $\gcd(x,1)=1$). This proves
\[\jacobi{-n}{p}=1 \iff p\mid x^2+ny^2,\ \gcd(x,y)=1.\]

Consider the conditions on $p$ in 1.1: The moduli for $n=1$ and 2 are already multiples of 4, and $p\equiv1\bmod 3$ is equivalent to $p\equiv 1, 4, 7, 10\bmod 12$. We know 4 and 10 are not coprime with 12 so we get $p\equiv 1, 7\bmod 12$ (and this agrees with Euler's conjecture about $(-3/p)$). It seems like the 4 essentially comes from $(-1/p)$ and is the smallest factor needed to write the conditions on $p$ in a unified way for all $n$.

\section{Computing $\jacobi{-1}{p}$ and $\jacobi{2}{p}$}

I took a small break to prove the formulas for $(-1/p)$ and $(2/p)$. These follow from Euler's criterion, which is proved like this:

Take $a \neq 0$ in $\mathbf{F}_p$ and let $g$ be a primitive element of $\mathbf{F}_p$. Suppose $(a/p) = 1$, then $a = g^{2i}$ for some $i$.
\[a^\frac{p-1}{2} = (g^{2i})^\frac{p-1}{2} = (g^{p-1})^i = 1.\]

Conversely, suppose $a^\frac{p-1}{2} = 1$ and take $j$ such that $a = g^j$. Then $g^{j\cdot\frac{p-1}{2}} = 1$, but the order of $g$ is $p-1$, so
\[\exists k,\ k(p-1)=j\cdot\frac{p-1}{2}\implies 2k=j.\]
Then $a = g^{2k} = (g^k)^2$ and $(a/p) = 1$. The cases $(a/p) = -1$ or 0 are immediate, which proves $(a/p) = a^\frac{p-1}{2}$. In particular, $a=-1$ yields
\[\jacobi{-1}{p} = (-1)^\frac{p-1}{2} =
  \begin{cases}
     1 &\quad p\equiv 1\bmod 4\\
    -1 &\quad p\equiv 3\bmod 4.
  \end{cases}
\]

For $a=2$, we can use the simple proof in Serre if we know that an algebraic closure (or splitting field) exists. For $a=3$, Serre's proof works almost exactly the same since if $\beta$ is a primitive 12th root of unity, $\beta^4-\beta^2+1=0$ so $\beta^2 + \beta^{-2} = 1$. Thus if $x=\beta + \beta^{-1}$, $x^2 = 3$ and the proof procedes the same way.

Here is an elementary proof for $a=2$: Fix a positive odd number $n = 2k+1$ and denote $\{i\in\mathbf{N} : 1\leq i\leq k\}$ by $[k]$.
Define $f:[k]\to[k]$ by
\[f(i) =
  \begin{cases}
    \frac{i}{2} &i\textrm{ even}\\
    \frac{n-i}{2} &i\textrm{ odd}.
  \end{cases}
\]
$f$ is surjective since
\begin{align*}
  i\leq k/2 &\implies 2i\leq k\textrm{ and }f(2i) = i\\
  i > k/2 &\implies n-2i < n-k = k+1\textrm{ and }f(n-2i) = i.
\end{align*}
Since the domain and codomain are equal, $f$ is a bijection. Thus,
\[\prod_{i=1}^k f(i) = k!\quad\implies\quad\prod_{i=1}^k (-1)^i\cdot 2\cdot f(i) = (-1)^\frac{k^2 + k}{2}\cdot 2^k\cdot k!.\]
Finally, $(-1)^i\cdot 2\cdot f(i)\equiv i\bmod n$, so
\[(-1)^\frac{k^2 + k}{2}\cdot 2^k\cdot k!\equiv k!\bmod n.\]
When $n$ is prime, we can cancel $k!$ and substitute $k=\frac{n-1}{2}$ to get
\[2^\frac{n-1}{2}\equiv (-1)^\frac{n^2 - 1}{8}\bmod n\]
which proves
\[\left(\frac{2}{n}\right) =
  \begin{cases}
     1 &\quad n\equiv\pm 1\bmod 8\\
    -1 &\quad n\equiv\pm 3\bmod 8.
  \end{cases}
\]

It looks like there are some group theoretic methods which are similar to Serre's method but without using splitting fields. Look into this more.

\section{$\jacobi{p^*}{q}=\jacobi{q}{p}\iff \jacobi{p}{q}\jacobi{q}{p}=(-1)^{(p-1)(q-1)/4}$}

For an odd prime $p$, define $p^*$ to be $(-1)^\frac{p-1}{2} p$. In other words,
\[p^*=
\begin{cases}
  p&p\equiv 1\bmod 4\\
  -p&p\equiv 3\bmod 4.\\
\end{cases}
\]

Assume quadratic reciprocity, and consider $(p^*/q)$ for distinct primes $p, q$. If $p\equiv 1\bmod 4$,
\[\jacobi{p^*}{q} = \jacobi{p}{q} = (-1)^{(q-1)\cdot\frac{(p-1)}{4}} = \jacobi{q}{p}.\]
If $p\equiv 3\bmod 4$,
\begin{align*}
  \jacobi{p^*}{q} &= \jacobi{-p}{q}\\
                  &= \jacobi{-1}{q}\jacobi{p}{q}\\
                  &= \jacobi{-1}{q}(-1)^{\frac{q-1}{2}\cdot\frac{p-1}{2}}\jacobi{q}{p}\\
                  &= \jacobi{-1}{q}\jacobi{-1}{q}^\frac{p-1}{2}\jacobi{q}{p}\\
                  &= \jacobi{-1}{q}^{1+\frac{p-1}{2}}\jacobi{q}{p} = \jacobi{q}{p}
\end{align*}
since $1+\frac{p-1}{2}$ is even when $p\equiv 3\bmod 4$. In both cases $(p^*/q)=(q/p)$.

Conversely, assume $(p^*/q) = (q/p)$. Then
\begin{align*}
  \jacobi{p}{q}\jacobi{q}{p} &= \jacobi{p}{q}\jacobi{p^*}{q}\\
                             &= \jacobi{p}{q}\jacobi{p}{q}\jacobi{-1}{q}^\frac{p-1}{2}\\
                             &= 1\cdot\jacobi{-1}{q}^\frac{p-1}{2}\\
                             &= \left((-1)^\frac{q-1}{2}\right)^\frac{p-1}{2}\\
                             &= (-1)^{(p-1)(q-1)/4}
\end{align*}
which is the usual statement of quadratic reciprocity.

This shows the equivalence of $(p^*/q) = (q/p)$ and quadratic reciprocity.

\section{$\jacobi{p^*}{q} = 1 \iff\exists b$ odd s.t. $p\equiv\pm b^2\bmod 4q$}

If $(p^*/q) = 1$, there exists $b$ such that $p^*\equiv b^2\bmod q$. We can assume $b$ is odd by adding $q$ if necessary. Consequently, $b^2\equiv 1\bmod 4$.

If $p\equiv 1\bmod 4$, $p\equiv b^2\bmod q\implies q\mid p-b^2$. Together with $4\mid p-b^2$, we have $p\equiv b^2\bmod 4q$.


If $p\equiv 3\bmod 4$, $-p\equiv b^2\bmod q\implies q\mid p+b^2$. Together with $4\mid p+b^2$, we have $p\equiv -b^2\bmod 4q$.

Conversely, suppose there is $b$ odd such that $p=b^2\bmod 4q$. Then
\[p=b^2\bmod 4\implies p=1\bmod 4\implies p^*=p,\textrm{ and}\]
\[p=b^2\bmod q\implies\jacobi{p}{q}=1\implies\jacobi{p^*}{q}=1.\]
If there is $b$ odd such that $p=-b^2\bmod 4q$, then
\[p=-b^2\bmod 4\implies p=3\bmod 4\implies p^*=-p,\textrm{ and}\]
\[p=-b^2\bmod q\implies\jacobi{-p}{q}=1\implies\jacobi{p^*}{q}=1.\]

\section{The Jacobi symbol as a character}

Quadratic reciprocity for the Jacobi symbol can be written
\[\left(\frac{n}{m}\right) = (-1)^{\frac{n-1}{2}\cdot\frac{m-1}{2}} \left(\frac{m}{n}\right).\]
I think it takes this form to allow for both sides to be zero.

In the following, we fix a nonzero integer $D\equiv 0, 1\bmod 4$.

Define a map
\[\chi:(\mathbf{Z}/D\mathbf{Z})^* \mapsto \{\pm 1\}\]
by the rule $[i]\mapsto (D/i)$ if $i$ is a positive odd integer.

First we show that the map is well-defined. We need the following lemma.

Lemma: If $k\equiv 0,1\bmod 4$ and $i,j$ are positive odd integers st $i\equiv j\bmod k$, then $(\frac{k}{i}) = (\frac{k}{j})$.
Proof: If $k\equiv 1\bmod 4$,
          Case: $k>0$. $\frac{k-1}{2}$ is even, so quadratic reciprocity and $(\frac{i}{k}) = (\frac{j}{k})$ yield
          $(\frac{k}{i}) = (\frac{i}{k}) = (\frac{j}{k}) = (\frac{k}{j})$.
          Case: $k<0$. Using $(\frac{k}{i}) = (\frac{-1}{i})(\frac{-k}{i})$ and quadratic reciprocity we get
          $(\frac{k}{i}) = (-1)^{\frac{i-1}{2}\cdot(1-\frac{k+1}{2}} (j/-k)$
                $= (j/-k)$ since $1-\frac{k+1}{2}$ is even.
          The same reasoning shows $(j/-k) = (\frac{k}{j})$.
       If $k\equiv 0\bmod 4$ and $k\neq 0$, write $k = 2^n \sgn(k) k^\prime$  where
       $2 \nmid k^\prime$ , $k^\prime >0$.
       $(\frac{k}{i}) = (\frac{2^n \sgn(k) k^\prime}{i})
             = (\frac{2}{i})^n (\frac{\sgn(k)}{i}) (\frac{k^\prime }{i}). Quadratic reciprocity and (\frac{i}{k^\prime}) = (\frac{j}{k^\prime})$ give
             $= (\frac{2}{i})^n (\frac{\sgn(k)}{i}) (-1)^{\frac{k^\prime-1)}{2}\left(\frac{i+j}{2} - 1\right)} (\frac{k^\prime }{j})$
       But since $i = j + 4c$ for some integer $c$ (remember $k\equiv 0\bmod 4$),
       $(i+j)/2 - 1 = (2j + 4c)/2 - 1 = j - 1 + 2c$ which is even. Thus,
             $= (\frac{2}{i})^n (\frac{\sgn(k)}{i}) (\frac{k^\prime }{j})
             = (\frac{2}{i})^n (\frac{\sgn(k)}{i}) (\frac{\sgn(k)}{j}) (\sgn(k)k^\prime /j)
             = (\frac{2}{i})^n (\sgn(k)k^\prime /j) since i\equiv j\bmod 4$.
       Finally, to put the powers of two back, we write $i = j + m*k^\prime *2^n$ and consider the following cases.
          Case: $n>2$. We have $i-j$ divisible by 8, so $(\frac{2}{i}) = (\frac{2}{j})$.
          Case: $n=2$. $(\frac{2}{i})^2 = (\frac{2}{j})^2$ is immediate.
          In both cases, $(\frac{2}{i})^n = (\frac{2}{j})^n$, so we can continue
       $(\frac{k}{i}) = (\frac{2}{j})^n (\sgn(k)k^\prime /j)
             = (2^n \sgn(k) k^\prime /j)
             = (\frac{k}{j})$.
       If $k=0$, the result is trivial, although we do not need this case.

(NOTE: In order to make this more clean, this lemma should be put with proofs of the other Jacobi symbol properties)

Continuing with well-definedness, suppose $[i] = [j]$ in $(\mathbf{Z}/D\mathbf{Z})^*$, $i$ and $j$ odd. $\chi([i]) = (\frac{D}{i}) and \chi([j]) = (\frac{D}{j})$. Then $(\frac{D}{i}) = (\frac{D}{j})$ is a direct application of the lemma.

Next we show that $\chi$ is total; specifically, for every $[i]\in (\mathbf{Z}/D\mathbf{Z})^*$, there is some positive odd $j$ such that $[j] = [i]$.
  Case: $i < 0$. $i+(-i)|D|$ and $i+(-i+1)|D|$ are both nonnegative and are in the same class as $[i]$. If $D$ is odd, they have opposite parity so let $j$ be the odd one. If $D$ is even, $i$ must be odd since $\gcd(i,D)=1$, so $j$ can be either one.
  Case: $i > 0$. Do the same as the previous case but with candidates $i+|D|$ and $i+2|D|$.

The map $\chi$ is a homomorphism since, assuming $a,b$ odd, $\chi([a][b]) = \chi([ab]) = (D/ab) = (\frac{D}{a})\cdot(\frac{D}{b}) = \chi([a])\cdot\chi([b])$ using a basic property of Jacobi symbols.

Now suppose we have a (possibly different) homomorphism
$\phi:(\mathbf{Z}/D\mathbf{Z})^* \mapsto {\pm 1}$, subject to the constraint that for any positive odd prime $p$ not dividing $D$ we have
  $\phi([p]) = (\frac{D}{p})$.

For any $[i]\in (\mathbf{Z}/D\mathbf{Z})^*$, we showed that there is a positive odd integer $j$ such that $[i] = [j]$. This means we can write $j=p_1^{a_1}\cdot\ldots\cdot p_n^{a_n}$ where each $p_i$ is a positive odd prime.

\begin{align*}
\textrm{Then }
  y([j]) &= y([p_1^{a_1}\cdot\ldots\cdot p_n^{a_n}])\\
         &= y([p_1])^{a_1}\cdot\ldots\cdot y([p_n])^{a_n}\\
         &= (D/p_1)^{a_1}\cdot\ldots\cdot (D/p_n)^{a_n}\\
         &= (D/p_1^{a_1}\cdot\ldots\cdot p_n^{a_n})\\
         &= (\frac{D}{j}) = \chi([j]).\\
\end{align*}

Thus $\chi$ is the \emph{unique} homomorphism $(\mathbf{Z}/D\mathbf{Z})^* \mapsto {\pm 1}$ such that $\chi([p]) = (\frac{D}{p})$ for positive odd primes $p\nmid D$.

We can calculate some values of $\chi$ explicitly.

Specifically, for $\chi([-1])$:
If $D>0$ and $D\equiv 1\bmod 4$ then $2D-1$ is odd and positive. Then
  \[\chi([-1]) = \chi([2D-1]) = (D/(2D-1)) = ((2D-1)/D) = (\frac{-1}{D}) = 1.\]
If $D>0$ and $D\equiv 0\bmod 4$ then $D-1$ is odd and positive. Then
  \[\chi([-1]) = \chi([D-1]) = (D/D-1) = (1/D-1) = 1.\]
If $D<0$ and $D\equiv 1\bmod 4$ then $-2D-1$ is odd and positive. Then
\begin{align*}
  \chi([-1]) &= \chi([-2D-1])\\
             &= (D/(-2D-1))\\
             &= (-D/(-2D-1))(-1/(-2D-1))\\
             &= (\frac{-1}{D})^*(-1)^((D+1)(D+1)/2) (-1/(-2D-1))\\
             &= -1 * 1 * 1 = -1.\\
\end{align*}
If $D<0$ and $D\equiv 0\bmod 4$ then $-D-1$ is odd and positive. Then
  \[(D/(-D-1)) = (-1/(-D-1))(-D/(-D-1)) = (-1)(1/(-D-1)) = -1.\]
Summarizing,
\[\chi([-1]) =
  \begin{cases}
    1&\quad D>0\\
    -1&\quad D<0.
  \end{cases}
\]

For $\chi([2])$, since 2 is even we must have $D$ odd (so $D\equiv 1\bmod 4$):
If $D>0$, $2+D$ is odd and positive.
  \[(D/2+D) = (2+D/D) = (\frac{2}{D}) = 1, D\equiv 1\bmod 8\textrm{ or }-1, D\equiv 5\bmod 8.\]
If $D<0$, $2-D$ is odd and positive.
  \[(D/2-D) = (2/2-D) = 1, 2-D=1\bmod 8 (\implies D\equiv 1\bmod 8)\textrm{ or }-1, 2-D\equiv 5\bmod 8 (\implies D\equiv 5\bmod 8).\]
Summarizing,
\[\chi([2]) =
  \begin{cases}
    1&\quad D\equiv 1\bmod 8\\
    -1&\quad D\equiv 5\bmod 8.
  \end{cases}
\]

One benefit of this homomorphism $\chi$ is that it gives an equivalent way of thinking about the reciprocity step in representing $x^2 + ny^2$.
Remember $p\mid x^2+ny^2$, $\gcd(x,y)=1 \iff (\frac{-n}{p})=1$.
This is almost the map $\chi$, except there are no restrictions on $n$ whereas $\chi$ needs the modulus to be 0 or $1\bmod 4$.
The trick is to notice that $(\frac{-n}{p})=1 \iff (-4n/p)=1$; now the set of primes $p$ dividing $x^2+ny^2$ are exactly the equivalence classes $[p]$ for which $\chi([p])=1$ (with $D=-4n$).

The book then gives a characterization of $\ker(x)$ which I don't see the significance of (besides having historical context) but maybe this will change as I read more.

\end{document}
